A longitudinal dataset of childhood gut microbiome linked with cognitive evaluation

### 2021-05-07 manuscript/main.tex
additional vanja edits

Here, we collected data on microbial taxonomic and functional composition
of the childhood microbiome paired with demographic information about participants
and age-appropriate cognitive assessments that produce IQ-like measures of motor, language, and social development.
Historically, many studies investigating the microbiome have used amplicon sequencing of conserved genes (eg 16S rRNA),
while more recently there has been a shift towards using shotgun metagenomic sequencing
which can provide more granular taxonomic assignment, including identification of fungal and viral components,
and direct evidence for gene functional potential..
Here, we provide both amplicon (16S rRNA V4-V5) and shotgun metagenomic sequencing
to enable the greatest potential for comparisons with previous studies
as well as the higher resolution enabled by sequencing the entire metagenome.
Understanding how the gut microbiome of healthy children
interacts with the complex metabolism of these and other neuroactive molecules
will be critical in understanding the etiology of neurocognitive disorders
and how to promote healthy neurological development.

@@

Data used in this study were drawn from the ongoing longitudinal RESONANCE study
of healthy and neurotypical brain and cognitive development,
based at Brown University in Providence, RI, USA.
RESONANCE is part of the NIH initiative Environmental influences on Child Health Outcomes (ECHO) \cite{Forrest2018-ud,Gillman2018-om},
a longitudinal observational study of healthy and neurotypical brain development
that spans the fetal and infant to adolescent life stages,
combining neuroimaging (magnetic resonance imaging, MRI), neurocognitive assessments, bio-specimen analyses, subject genetics,
environmental exposures such as lead, and rich demographic, socioeconomic, family and medical history information.
From the RESONANCE cohort, 344 typically-developing children
between the ages of 28 days and 15 years oldwere selected for analysis in this study.

@@

Stool samples (n=493) were collected by parents in OMR-200 tubes (OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada).
immediately stored on ice, and brought within 24 hrs to the lab in RI where they were immediately frozen at -80$^{\circ}$C.
Stool samples were not collected if the subject had taken antibiotics within the last two weeks.
DNA extraction was performed at Wellesley College (Wellesley, MA).
Nucleic acids were extracted from stool samples using the RNeasy PowerMicrobiome kit
automated on the QIAcube (Qiagen, Germantown, MD), excluding the DNA degradation steps.
Extracted DNA was sequenced at the Integrated Microbiome Resource (IMR, Dalhousie University, NS, Canada)

@@

Data records, other than biological sequence data, can be accessed from the Open Science Foundation (OSF.io) repository \cite{Bonham2021-zy},
or using the \verb|ResonanceMicrobiome.jl| package in a julia script or interactive session.
Instructions for using the \verb|ResonanceMicrobiome.jl| package can be found in the README in the code repository on github.com (https://github.com/Klepac-Ceraj-Lab/ResonanceMicrobiome).

@@

Amplicon sequence variants (ASV) obtained from DADA2 and mapping to SILVA taxonomy are available on OSF.io
in the “amplicon-feature-table.tsv” and “amplicon-taxonomy.tsv” files respectively,
and can be accessed in a julia session with \verb|ResonanceMicrobiome.jl| using \verb|datadep”amplicon_features”|,
and \verb|datadep”amplicon_taxonomy”| respectively.

@@

Code for downloading data records (excluding raw sequence files) and generating included figures and tables,
as well as convenience functions for interacting with the data in a julia interactive session,
are available on github.
Instructions for installing julia and this package code are available in the README for the repository.
Releases made for initial submissions, revisions, and the final manuscript
are also available on github and archived on Zenodo \cite{Bonham2021-gl}

>>>>

Here, we collected data on microbial taxonomic and functional composition
of the childhood microbiome paired with demographic information about participants
and age-appropriate cognitive assessments that produce IQ-like measures of motor, language, and social development.
Historically, many studies investigating the microbiome have used amplicon sequencing of conserved genes (e.g., 16S rRNA),
while more recently there has been a shift towards using shotgun metagenomic sequencing
which can provide more granular taxonomic assignment, including identification of fungal and viral components,
and direct evidence for gene functional potential.
Here, we provide both amplicon (16S rRNA V4-V5) and shotgun metagenomic sequencing
to enable the greatest potential for comparisons with previous studies
as well as the higher resolution enabled by sequencing the entire metagenome.
Understanding how the gut microbiome of healthy children
interacts with the complex metabolism of these and other neuroactive molecules
will be critical in understanding the etiology of neurocognitive disorders
and how to promote healthy neurological development.

@@

Data used in this study were drawn from the ongoing longitudinal RESONANCE study
of healthy and neurotypical brain and cognitive development,
based at Brown University in Providence, RI, USA.
RESONANCE is part of the NIH initiative Environmental influences on Child Health Outcomes (ECHO) \cite{Forrest2018-ud,Gillman2018-om},
a longitudinal observational study of healthy and neurotypical brain development
that spans the fetal and infant to adolescent life stages,
combining neuroimaging (magnetic resonance imaging, MRI), neurocognitive assessments, bio-specimen analyses, subject genetics,
environmental exposures such as lead, and rich demographic, socioeconomic, family and medical history information.
From the RESONANCE cohort, 344 typically-developing children
between the ages of 28 days and 15 years old were selected for analysis in this study.

@@

Stool samples (n=493) were collected by parents in OMR-200 tubes (OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada),
immediately stored on ice, and brought within 24 hrs to the lab in RI where they were immediately frozen at -80 $^{\circ}$C.
Stool samples were not collected if the subject had taken antibiotics within the last two weeks.
DNA extraction was performed at Wellesley College (Wellesley, MA).
Nucleic acids were extracted from stool samples using the RNeasy PowerMicrobiome kit
automated on the QIAcube (Qiagen, Germantown, MD), excluding the DNA degradation steps.
Extracted DNA was sequenced at the Integrated Microbiome Resource (IMR, Dalhousie University, NS, Canada)

@@

Data records, other than biological sequence data, can be accessed from the Open Science Foundation (OSF.io) repository \cite{Bonham2021-zy},
or using the \verb|ResonanceMicrobiome.jl| package in a julia script or interactive session.
Instructions for using the \verb|ResonanceMicrobiome.jl| package can be found in the README in the code repository at https://github.com/Klepac-Ceraj-Lab/ResonanceMicrobiome.

@@

Amplicon sequence variants (ASV) obtained from DADA2 and mapping to SILVA taxonomy are available on OSF.io
in the “amplicon-feature-table.tsv” and “amplicon-taxonomy.tsv” files respectively,
and can be accessed in a julia session with \verb|ResonanceMicrobiome.jl| using \verb|datadep”amplicon_features,”|
and \verb|datadep”amplicon_taxonomy”| respectively.

@@

Code for downloading data records (excluding raw sequence files) and generating included figures and tables,
as well as convenience functions for interacting with the data in a julia interactive session,
are available on github.
Instructions for installing julia and this package code are available in the README for the repository.
Releases made for initial submissions, revisions, and the final manuscript
are also available on github and archived on Zenodo \cite{Bonham2021-gl}.
